A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-16 ionic_liquids.tex
Initial revision of the concentrated CO2 solutions section

\section{Introduction}

@@

Three additional assumptions are made here. First, we assume that the gas phase is either pure solute (which is consistent with the ionic-liquid systems with which we compare our simulations) or can be approximated by an ideal solution. Second, we assume that the system pressure is moderate, so that a truncated virial equation is enough to represent the gas phase. In this case, Eq.~\eqref{eq:fgas_d} becomes
\begin{equation}
\label{eq:equation for pressure}
y_s \exp\left( \frac{B_s P}{R T} \right) P = \exp \left( \frac{\Delta G_\text{sim}}{R T} \right) +  c_s R T,
\end{equation}
where $y_s$ is the molar fraction of solute in the gas phase ($y_s \rightarrow 1$ for the systems we analyze here) and $B_s$ is the second virial coefficient of the solute.
In the case of \ce{CO_2}, we employ the correlation for $B_s$ developed by Holste \textit{et al}. \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0 = 23.02991~\mathrm{\frac{cm^3}{mol}}$, $B_1 = -2455.297~\mathrm{\frac{K \cdot cm^3}{mol}}$, and $B_2 = -1.22675 \times 10^7~\mathrm{\frac{K^2 \cdot cm^3}{mol}}$.

Our third assumption, supported by preliminary simulations, is that $\Delta G_\text{sim}$ does not vary appreciably with pressure, at least over the range of experimental data considered here.
In this way, no iterative simulations are required to compute the equilibrium pressure when solving Eq.~\eqref{eq:equation for pressure} for $P$.

@@

Since concentrated solutions are the more likely scenario in \ce{CO_2} capture processes, we have also carried out simulations at the thermodynamic conditions specified in Table~\ref{table:solv}, which correspond to similar values of \ce{CO_2} molality.
Note that the differences in the \ce{CO_2} molar fractions clearly exemplifies the observation made by Carvalho \textit{et al}. \cite{Carvalho_2016}, mentioned at the introduction of this paper.

\begin{table}
	\centering
	\caption{Thermodynamic conditions employed in the solvation study at high concentrations of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}~\cite{Makino_2014,Schilderman_2007}.}
	\begin{tabular}{ c  c  c  }  
		\hline\hline
		IL & \ce{[emim][B(CN)_4]}  & \ce{[emim][NTf_2]} \\
		\hline
		model &  Ref.~\citenum{Koller_2012} & Ref.~\citenum{Koddermann_2007} \\
		T  & 313.15 K & 312.13 K \\
		P  & 21 atm & 37 atm \\
		$x_s$ & 0.352 & 0.479 \\
		$m_s$ & 2.4 mol/kg & 2.35 mol/kg \\
		\hline\hline
		\label{table:solv} 
	\end{tabular}
\end{table}

For each system, in Fig.~\ref{fig:pressure} we show our predictions of the gas-phase pressure ($P_s$), as described in Sec.~\ref{sec:partial_pressure}.
The dotted lines are the experimental pressures, which correspond to the specified values ($P$) in the liquid-phase simulation.
Note that our results reproduce what is experimentally observed, in the sense that the calculated $P_s$ is considerably lower for \ce{CO_2}-\ce{[emim][B(CN)_4]}.
In addition, we see that in this case the correction to the ideal gas assumption is almost negligible, as $P_s$ is very close to the fugacity ($f^g_s$).
The solvation free energies obtained for \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-\ce{[emim][NTf_2]} were $-0.755 \pm 0.01 ~\mathrm{kcal/mol}$ and $-0.542 \pm 0.005$, respectively.
However, since we simulated a different number of \ce{CO_2}, we cannot draw any definite conclusion regarding the presence of stronger interactions with \ce{[emim][B(CN)_4]}, which could explain the higher solubility.

>>>>

\section{Introduction}
\label{sec:intro}

@@

Three additional assumptions are made here. First, we assume that the gas phase is either pure solute (which is consistent with the ionic-liquid systems with which we compare our simulations) or can be approximated by an ideal solution. Second, we assume that the system pressure is moderate, so that a truncated virial equation is enough to represent the gas phase. In this case, Eq.~\eqref{eq:fgas_d} becomes
\begin{equation}
\label{eq:equilibrium pressure}
y_s \exp\left( \frac{B_s P}{R T} \right) P = \exp \left( \frac{\Delta G_\text{sim}}{R T} \right) +  c_s R T,
\end{equation}
where $y_s$ is the molar fraction of solute in the gas phase ($y_s \rightarrow 1$ for the systems we analyze here) and $B_s$ is the second virial coefficient of the solute.
In the case of \ce{CO_2}, we employ the correlation for $B_s$ developed by Holste \textit{et al}. \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0 = 23.02991~\mathrm{\frac{cm^3}{mol}}$, $B_1 = -2455.297~\mathrm{\frac{K \cdot cm^3}{mol}}$, and $B_2 = -1.22675 \times 10^7~\mathrm{\frac{K^2 \cdot cm^3}{mol}}$.

Our third assumption, supported by preliminary simulations, is that $\Delta G_\text{sim}$ does not vary appreciably with pressure, at least over the range of experimental data considered here.
In this way, no iterative simulations are required to compute the equilibrium pressure when solving Eq.~\eqref{eq:equilibrium pressure} for $P$.

@@

Since concentrated solutions are the more likely scenario in carbon capture processes, we have also carried out simulations of systems with higher \ce{CO_2} concentrations.
The thermodynamic conditions specified for these simulations, which are based on experimental investigations \cite{Makino_2014,Schilderman_2007}, are shown in Table~\ref{table:solv}.
We have selected these specific conditions because of the proximity observed in the molality of \ce{CO_2} ($m_s$).
It is worth noting that the marked difference in the molar fractions of \ce{CO_2} in Table~\ref{table:solv} does not imply that \ce{[emim][NTf_2]} is a better solvent than \ce{[emim][B(CN)_4]}, thus exemplifying the observation made by Carvalho \textit{et al}. \cite{Carvalho_2016} and discussed here in Sec.~\ref{sec:intro}.

\begin{table}
	\centering
	\caption{Thermodynamic conditions employed in the solvation study at high concentrations of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}, based on the experimental observations of Refs.~ \citenum{Makino_2014} and \citenum{Schilderman_2007}.}
	\begin{tabular}{ccc}
		\hline\hline
		IL & \ce{[emim][B(CN)_4]}  & \ce{[emim][NTf_2]} \\
		\hline
%		model & Ref.~\citenum{Koller_2012} & Ref.~\citenum{Koddermann_2007} \\
		T  & 313.15 K & 312.13 K \\
		P  & 21 atm & 37 atm \\
		$x_s$ & 0.352 & 0.479 \\
		$m_s$ & 2.4 mol/kg & 2.35 mol/kg \\
		\hline\hline
		\label{table:solv} 
	\end{tabular}
\end{table}

We carried out expanded ensemble simulations of liquid solutions at the conditions of Table~\ref{table:solv}.
The model of Koller \textit{et al}. \cite{Koller_2012} was the one employed for \ce{[emim][B(CN)_4]}.
The computed solvation free energies for the systems \ce{CO_2}+\ce{[emim][B(CN)_4]} and \ce{CO_2}+\ce{[emim][NTf_2]} were $-0.755 \pm 0.01 ~\mathrm{kcal/mol}$ and $-0.542 \pm 0.005~\mathrm{kcal/mol}$, respectively.
In Fig.~\ref{fig:pressure}, we present our predictions of the equilibrium pressure ($P$) obtained by solving Eq.~\eqref{eq:equilibrium pressure}, as well as the corresponding gas-phase fugacities of \ce{CO_2} ($f_s^g$) computed by Eq.~\eqref{eq:fgas_d}.
The dotted lines correspond to the experimental values (i.e. the ones used to simulate the liquid solutions).
First, the results make clear that assuming ideal gas would have been an oversimplification, specially in the case of \ce{[emim][NTf_2]}.
Second, the computed pressures did not depart considerably from the specified values, meaning that our results sustain the experimental observation that $P$ is considerably lower for the system \ce{CO_2}+\ce{[emim][B(CN)_4]} than for the system \ce{CO_2}+\ce{[emim][NTf_2]}.
In addition, we see that in this case the correction to the ideal gas assumption is almost negligible, as $P_s$ is very close to the fugacity ($f^g_s$).
%However, since we simulated a different number of \ce{CO_2}, we cannot draw any definite conclusion regarding the presence of stronger interactions with \ce{[emim][B(CN)_4]}, which could explain the higher solubility.
